---
title: "Temporal analysis of rounds 8-13 of REACT-1 when national restrictions were eased"
author: "Oliver Eales"
date: "12 September, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load reactidd}
library(reactidd)
```

# Loading the Data
First we will load the REACT data

```{r loading the react data}
pos <- load_example_data_weighted()[[1]]
tot <- load_example_data_weighted()[[2]]

```

In order to subset the data we define the dates for rounds 14 to 18 of the study
```{r setting the minimum and maximum dates}
min_date_r1 <- as.Date("2020-05-01")
min_date_r8 <- as.Date("2020-12-30")
max_date_r13 <- as.Date("2021-07-12")
```



# Fitting the Bayesian P-Spline Model to the data
We can then fit the Bayesian P-spline model to all of the REACT data for rounds 1 to 13 of the study.

```{r Fitting the p-spline model, results='hide'}

p_spline_mod_react <- stan_p_spline_weighted(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   iter = 2000,
                                   warmup = 500,
                                   cores = 1)


```

The p-spline model can then be plotted with estimated 95%CI  for prevalence (all the raw data used to produce the plots is also avialble in the list returned). In the paper this analysis was perfomed for we only present the model for rounds 8-13.
```{r Plotting the p-spline model}
p_spline_plot <- plot_p_spline_prev(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   p_spline_fit = p_spline_mod_react, 
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   ylim = 3.0)

print(p_spline_plot[[1]])

```


We can then calculate and plot the rolling two week average Reproduction number using a generation time suitable for Alpha 
```{r Reproduction number calculation}
p_spline_Rt_A <- plot_p_spline_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   ylim = 2.,
                                   n1=2.29, b1=0.36,
                                   tau_max = 14)

print(p_spline_Rt_A[[1]])


```


Or as a sensitivity analysis using a generation time approporate for the Delta variant.
```{r Reproduction number calculation}
p_spline_Rt_D <- plot_p_spline_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   ylim = 2.,
                                   n1=2.20, b1=0.48,
                                   tau_max = 14)

print(p_spline_Rt_D[[1]])


```


# Fitting the Segmented model to the data
We can fit the Segmented model to all of the REACT data for rounds 8 to 13 of the study.

```{r Fitting the segmented model, results='hide'}

#Setting the key dtaes of the changes in restrictions
res0 <- as.Date("2021-01-06")
res1 <- as.Date("2021-03-08")
res2 <- as.Date("2021-03-29")
res3 <- as.Date("2021-04-12")
res4 <- as.Date("2021-05-17")

rest_dates <- c(res0,res1,res2,res3,res4)

breakpoint_mod_react <- stan_breakpoint_model_weighted(pos[pos$X>=min_date_r8 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r8 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r8 & tot$X<= max_date_r13,]$England,
                                   restriction_dates = rest_dates,
                                   iter = 2000,
                                   warmup = 500,
                                   cores = 1)


```


We can then plot the estimated R's and multiplicative changes in R as inferred from the Segmented model

```{r Plotting the segmented model, results='hide'}


```



# Fitting the Segmented model with vaccine + variant effects
We can fit the Segmented model including effects due to vaccination and the Delta variant to all of the REACT data for rounds 8 to 13 of the study.


```{r Read in data on vaccination and variant proportion, results='hide'}




```

```{r Fitting the complex segmented model, results='hide'}

breakpoint_mod_complex_react <- stan_breakpoint_model_complex_weighted(pos[pos$X>=min_date_r8 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r8 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r8 & tot$X<= max_date_r13,]$England,
                                   restriction_dates = rest_dates,
                                   iter = 2000,
                                   warmup = 500,
                                   cores = 1)


```


We can then plot the estimated R's and multiplicative changes in R as inferred from the Segmented model

```{r Plotting the complex segmented model, results='hide'}


```
