---
title: "Temporal analysis of rounds 8-13 of REACT-1 when national restrictions were eased"
author: "Oliver Eales"
date: "12 September, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load reactidd}
library(reactidd)
```

# Loading the Data
First we will load the REACT data

```{r loading the react data}
pos <- load_example_data_weighted()[[1]]
tot <- load_example_data_weighted()[[2]]

```

In order to subset the data we define the dates for rounds 14 to 18 of the study
```{r setting the minimum and maximum dates}
min_date_r1 <- as.Date("2020-05-01")
min_date_r8 <- as.Date("2020-12-30")
max_date_r13 <- as.Date("2021-07-12")
```



# Fitting the Bayesian P-Spline Model to the data
We can then fit the Bayesian P-spline model to all of the REACT data for rounds 1 to 13 of the study.

```{r Fitting the p-spline model, results='hide'}

p_spline_mod_react <- stan_p_spline_weighted(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   iter = 2000,
                                   warmup = 500,
                                   cores = 1)


```

The p-spline model can then be plotted with estimated 95%CI  for prevalence (all the raw data used to produce the plots is also avialble in the list returned). In the paper this analysis was perfomed for we only present the model for rounds 8-13.
```{r Plotting the p-spline model}
p_spline_plot <- plot_p_spline_prev(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   p_spline_fit = p_spline_mod_react, 
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   ylim = 3.0)

print(p_spline_plot[[1]])

```


We can then calculate and plot the instanteous growth rate over the study period

```{r Instantaneous growth rate calculation}
p_spline_igr <- plot_p_spline_igr_two_variants(X = pos[pos$X>=min_date_r14 & pos$X<= max_date_r18,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   ylim = 0.5,
                                   labs =c("Delta","Omicron"),
                                   colors1 = RColorBrewer::brewer.pal(3, "Dark2"),
                                   mindateVar1 = as.Date("2021-09-09"), maxdateVar1 = as.Date("2022-02-14"),
                                   mindateVar2 = as.Date("2021-12-03"), maxdateVar2 = as.Date("2022-03-03"))

print(p_spline_igr[[1]])


```
and also the rolling two week average Reproduction number
```{r Reproduction number calculation}
p_spline_Rt <- plot_p_spline_R_two_variants(X = pos[pos$X>=min_date_r14 & pos$X<= max_date_r18,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   ylim = 2.5,
                                   n1=2.20, b1=0.48,
                                   n2=0.89, b2=0.27,
                                   tau_max = 14,
                                   labs =c("Delta","Omicron"),
                                   colors1 = RColorBrewer::brewer.pal(2, "Set1"),
                                   mindateVar1 = as.Date("2021-09-09"), maxdateVar1 = as.Date("2022-02-14"),
                                   mindateVar2 = as.Date("2021-12-03"), maxdateVar2 = as.Date("2022-03-03"))

print(p_spline_Rt[[1]])


```



