---
title: "Fitting models to REACT data"
author: "Oliver Eales"
date: "November 23, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load reactidd}
library(reactidd)
```

# Loading the Data
First we will load the example REACT data available using reactidd::load_example_data() 

```{r loading the react data}
pos <- load_example_data()[[1]]
tot <- load_example_data()[[2]]
```

In order to subset the data we define the dates of each round of the study
```{r setting the minimum and maximum dates}
min_date_r1 <- as.Date("2020-05-01")
min_date_r2 <- as.Date("2020-06-19")
min_date_r3 <- as.Date("2020-07-24")
min_date_r4 <- as.Date("2020-08-22")

max_date_r1 <- as.Date("2020-06-01")
max_date_r2 <- as.Date("2020-07-07")
max_date_r3 <- as.Date("2020-08-11")
max_date_r4 <- as.Date("2020-09-08")
```


# Fitting the exponential model
We fit the exponential model to subsets of the data corresponding to individual rounds and pairs of subsequent rounds 

```{r Fitting the models, results='hide'}
exp_mod_react_r1 <- stan_exp_model(pos[pos$X>=min_date_r1 & pos$X<= max_date_r1,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r1,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r1,]$England,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)
exp_mod_react_r2 <- stan_exp_model(pos[pos$X>=min_date_r2 & pos$X<= max_date_r2,]$X,
                                   pos[pos$X>=min_date_r2 & pos$X<= max_date_r2,]$England,
                                   tot[tot$X>=min_date_r2 & tot$X<= max_date_r2,]$England,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)
exp_mod_react_r3 <- stan_exp_model(pos[pos$X>=min_date_r3 & pos$X<= max_date_r3,]$X,
                                   pos[pos$X>=min_date_r3 & pos$X<= max_date_r3,]$England,
                                   tot[tot$X>=min_date_r3 & tot$X<= max_date_r3,]$England,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)
exp_mod_react_r4 <- stan_exp_model(pos[pos$X>=min_date_r4 & pos$X<= max_date_r4,]$X,
                                   pos[pos$X>=min_date_r4 & pos$X<= max_date_r4,]$England,
                                   tot[tot$X>=min_date_r4 & tot$X<= max_date_r4,]$England,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)
exp_mod_react_r12 <- stan_exp_model(pos[pos$X>=min_date_r1 & pos$X<= max_date_r2,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r2,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r2,]$England,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)
exp_mod_react_r23 <- stan_exp_model(pos[pos$X>=min_date_r2 & pos$X<= max_date_r3,]$X,
                                   pos[pos$X>=min_date_r2 & pos$X<= max_date_r3,]$England,
                                   tot[tot$X>=min_date_r2 & tot$X<= max_date_r3,]$England,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)
exp_mod_react_r34 <- stan_exp_model(pos[pos$X>=min_date_r3 & pos$X<= max_date_r4,]$X,
                                   pos[pos$X>=min_date_r3 & pos$X<= max_date_r4,]$England,
                                   tot[tot$X>=min_date_r3 & tot$X<= max_date_r4,]$England,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)
```

Using these model fits we can calculate the growth rate, R and the doubling/halving times for each model fit 
```{r Calculating R of models}
R_estimates_react_r1 <- exponential_estimate_R(exp_mod_react_r1, n_mean = 2.29, b_mean =0.36, label ="React-Round1")
R_estimates_react_r2 <- exponential_estimate_R(exp_mod_react_r2, n_mean = 2.29, b_mean =0.36, label ="React-Round2")
R_estimates_react_r3 <- exponential_estimate_R(exp_mod_react_r3, n_mean = 2.29, b_mean =0.36, label ="React-Round3")
R_estimates_react_r4 <- exponential_estimate_R(exp_mod_react_r4, n_mean = 2.29, b_mean =0.36, label ="React-Round4")
R_estimates_react_r12 <- exponential_estimate_R(exp_mod_react_r12, n_mean = 2.29, b_mean =0.36, label ="React-Round1&2")
R_estimates_react_r23 <- exponential_estimate_R(exp_mod_react_r23, n_mean = 2.29, b_mean =0.36, label ="React-Round2&3")
R_estimates_react_r34 <- exponential_estimate_R(exp_mod_react_r34, n_mean = 2.29, b_mean =0.36, label ="React-Round3&4")

R_table <- rbind(R_estimates_react_r1, R_estimates_react_r2, R_estimates_react_r3, R_estimates_react_r4,
                 R_estimates_react_r12, R_estimates_react_r23, R_estimates_react_r34)
print(R_table)
```

# Plotting the exponential model fits

The model fits can then be plotted. First the individual round fits:

```{r Plotting Individual rounds}
individual_round_plots <- plot_exp_model(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$X,
                                   Y= pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$England,
                                   N = tot[tot$X>=min_date_r1 & tot$X<= max_date_r4,]$England,
                                   fit_exp = list(exp_mod_react_r1, exp_mod_react_r2, exp_mod_react_r3, exp_mod_react_r4),
                                   X_model = list(pos[pos$X>=min_date_r1 & pos$X<= max_date_r1,]$X,
                                                  pos[pos$X>=min_date_r2 & pos$X<= max_date_r2,]$X,
                                                  pos[pos$X>=min_date_r3 & pos$X<= max_date_r3,]$X,
                                                  pos[pos$X>=min_date_r4 & pos$X<= max_date_r4,]$X),
                                   color_list = list("red","red","red","red"),
                                   ylim = 1.0)

print(individual_round_plots[[1]])
```

Then plot the models fit to subsequent rounds
```{r Plotting Subsequent rounds}
subsequent_round_plots <- plot_exp_model(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$X,
                                   Y= pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$England,
                                   N = tot[tot$X>=min_date_r1 & tot$X<= max_date_r4,]$England,
                                   fit_exp = list(exp_mod_react_r12, exp_mod_react_r23, exp_mod_react_r34),
                                   X_model = list(pos[pos$X>=min_date_r1 & pos$X<= max_date_r2,]$X,
                                                  pos[pos$X>=min_date_r2 & pos$X<= max_date_r3,]$X,
                                                  pos[pos$X>=min_date_r3 & pos$X<= max_date_r4,]$X),
                                   color_list = list("red","blue","dark green"),
                                   ylim = 1.0)

print(subsequent_round_plots[[1]])
```

# Fitting the Bayesian P-Spline Model to the data
We can then fit the Bayesian P-spline model to all of the REACT data for the first 4 rounds of the study

```{r Fitting the p-spline model, results='hide'}
p_spline_mod_react <- stan_p_spline(pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r4,]$England,
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   iter = 5000,
                                   warmup = 1000,
                                   cores = 1)


```

The p-spline model can then be plotted with estimated 95%CI and 50%CI
```{r Plotting the p-spline model}
p_spline_plot <- plot_p_spline_prev(pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r4,]$England,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   ylim = 1.0)

print(p_spline_plot[[1]])


```

From the p-spline model we can estimate the date of minimum prevalence and plot the posterior probability density

```{r Date of minimum from P-Spline model}
p_spline_min_date <- plot_p_spline_minimum_density(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r4,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3)

print(p_spline_min_date[[1]])


```

