---
title: "Fitting models to estimate duration of swab-test pcr-positivity"
author: "Oliver Eales"
date: "5 February, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load reactidd}
library(reactidd)
```

## Estimating duration of positivity and sensitivity
First we must simlate some indiviudal level data. We will simulate two datasets one using an exponential decay model and another with a time delay before exponential decay. Default parameters of the function are sensitivity=0.79 and decay rate = 0.071
```{r simualte individual level shedding data}
shed_dat1 <- shedding_simulate_data()
shed_dat2 <- shedding_simulate_data(tau=3.97, sens=0.95, k=0.064)
```

Next we will convert the individual data into a structure that can be input into the stan models

```{r reformat indiviudla level data}
dat1 <- shedding_prepare_data_for_binary_model(shed_dat1)
dat2 <- shedding_prepare_data_for_binary_model(shed_dat2)
```

# Fitting exponential decay models
We will now fit two models to each data set 1) the exponential decay model of positivity 2) the exponential decay model of positivity with a time delay.

```{r fit exp decay models}
mods1 <- stan_both_shedding_exp_models(dat1, iter=2000, warmup = 500, chains=1)
mods2 <- stan_both_shedding_exp_models(dat2, iter=2000, warmup = 500, chains=1)

```

We can compare models using the ELPD

```{r compare exp decay models}
library(loo)
loo_dat1_mod1<-loo::loo(mods1[[1]])
loo_dat1_mod2<-loo::loo(mods1[[2]])

loo_dat2_mod1<-loo::loo(mods2[[1]])
loo_dat2_mod2<-loo::loo(mods2[[2]])

loo::loo_compare(loo_dat1_mod1, loo_dat1_mod2)
loo::loo_compare(loo_dat2_mod1, loo_dat2_mod2)


```

or plot them 
```{r plot exp decay models}

plot1 <- shedding_plot_exp_models(shed_dat1, mods1[[1]], mods1[[2]], label_name = "Simulation1")
plot2 <- shedding_plot_exp_models(shed_dat2, mods2[[1]], mods2[[2]], label_name = "Simulation2")

print(plot1)
print(plot2)
```

We can extract the parameter values and credible intervals as well
```{r get parameter values from exp models}
row1 <- shedding_get_params(mods1[[1]], label="Data 1 Model 1")
row2 <- shedding_get_params(mods1[[2]], label="Data 1 Model 2")
row3 <- shedding_get_params(mods2[[1]], label="Data 2 Model 1")
row4 <- shedding_get_params(mods2[[2]], label="Data 2 Model 2")

tab <- rbind(row1, row2, row3, row4)
print(tab)

```

Or plot a 2d density map of the posterior

```{r plot 2d density maps}

plot1 <-shedding_exp_model_plot_density(mods1[[1]], label_name = "Data 1 Model 1")
plot2 <-shedding_exp_model_plot_density(mods1[[2]], label_name = "Data 1 Model 2")
plot3 <-shedding_exp_model_plot_density(mods2[[1]], label_name = "Data 2 Model 1")
plot4 <-shedding_exp_model_plot_density(mods2[[2]], label_name = "Data 2 Model 2")

print(plot1)
print(plot2)
print(plot3)
print(plot4)

```

## Caclulating Incidence and Swab-positivity over time
Now we will load the weighted REACT data available using reactidd::load_example_data_weighted() 
Datasets have the weighted number of positives by english region and for England as a whole (pos below)
and the weighted number of total tests by english region and for England as a whole (tot)

```{r loading the react data}
pos <- load_example_data_weighted()[[1]]
tot <- load_example_data_weighted()[[2]]
```

In order to subset the data we define the dates of each round of the study
```{r setting the minimum and maximum dates}
min_date_r1 <- as.Date("2020-05-01")
min_date_r13 <- as.Date("2021-06-24")

max_date_r1 <- as.Date("2020-06-01")
max_date_r13 <- as.Date("2021-07-12")
```



# Fitting the Bayesian P-Spline Model to the data
We can then fit the Bayesian P-spline model to all of the REACT data for the first 13 rounds of the study
Note that a larger number of iterations is required to get a reliable fit.

```{r Fitting the p-spline model, results='hide'}
p_spline_mod_inc <- stan_p_spline_weighted_incidence(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   iter = 2000,
                                   warmup = 500,
                                   cores = 1,
                                   chains=1)


```

```{r Fitting the p-spline model, results='hide'}
p_spline_mod_prev <- stan_p_spline_weighted(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   iter = 2000,
                                   warmup = 500,
                                   cores = 1,
                                   chain = 1)


```

The p-spline model can then be plotted with estimated 95% and 50% credible intervals
```{r Plotting the p-spline model}
p_spline_plot <- plot_p_spline_prev(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   p_spline_fit = p_spline_mod_prev, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   ylim = 3.0)

print(p_spline_plot[[1]])

p_spline_plot[[1]] <- p_spline_plot[[1]]+
  ggplot2::coord_cartesian(xlim=c(min_date_r1, max_date_r13), ylim = c(0.01,3))+
  ggplot2::scale_y_log10()

print(p_spline_plot[[1]])


```

```{r Plotting the p-spline model}
p_spline_inc_plot <- plot_p_spline_inc(pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r13,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r13,]$England,
                                   p_spline_fit = p_spline_mod_inc,
                                   prev_plot = p_spline_plot[[1]],
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   ylim = 0.3)




print(p_spline_inc_plot[[1]])


```



