---
title: "Fitting models to PHE pillar 1&2 data for the period of REACT-1 rounds 1-7"
author: "Oliver Eales"
date: "4 February, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load reactidd}
library(reactidd)
```

# Loading the Data
First we will load the example REACT data available using reactidd::load_example_data() 

```{r loading the react data}
phe <- load_example_phe_data()

```

In order to subset the data we define the dates  of the study
```{r setting the minimum and maximum dates}
min_date_r1 <- as.Date("2020-05-01")
max_date_r7 <- as.Date("2020-12-03")
```



# Fitting the Bayesian P-spline model
We fit the bayesian p-spline model to all the data corresponding to the period of time of the first 4 rounds of the REACT study 

```{r Fit the P-Spline model, echo=TRUE, results='hide'}
p_spline_mod_phe <- stan_p_spline_phe(X = phe[phe$date>=min_date_r1 & phe$date<= max_date_r7,]$date,
                                   Y= phe[phe$date>=min_date_r1 & phe$date<= max_date_r7,]$n_cases,
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   iter = 2000,
                                   warmup = 500,
                                   cores = 1)


```

We can then plot the model fit with 95%CI and 50%CI
```{r Plotting the p-spline model}

p_spline_plot <- plot_p_spline_phe(X = phe[phe$date>=min_date_r1 & phe$date<= max_date_r7,]$date,
                                   Y= phe[phe$date>=min_date_r1 & phe$date<= max_date_r7,]$n_cases,
                                   p_spline_fit = p_spline_mod_phe, 
                                    target_dist_between_knots = 5,
                                  spline_degree = 3,
                                   ylim = 5000.0)


print(p_spline_plot[[1]])


```

From the p-spline model we can estimate the date of minimum prevalence and plot the posterior distribution

```{r Date of minimum from P-Spline model}
p_spline_min_date <- plot_p_spline_minimum_density(X = phe[phe$date>=min_date_r1 & phe$date<= max_date_r4,]$date,
                                   p_spline_fit = p_spline_mod_phe, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3)

print(p_spline_min_date[[1]])
```
