---
title: "Fitting Omicron vs Delta P-spline to REACT-1 data for rounds 14-18"
author: "Oliver Eales"
date: "17 May, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load reactidd}
library(reactidd)
```

# Loading the Data
First we will load the REACT data available using reactidd::load_example_data() 

```{r loading the react data}
pos <- load_example_data_weighted()[[1]]
tot <- load_example_data_weighted()[[2]]
```

In order to subset the data we define the dates for rounds 14 to 18 of the study
```{r setting the minimum and maximum dates}
min_date_r14 <- as.Date("2021-09-09")
max_date_r18 <- as.Date("2022-03-04")
```



# Fitting the Bayesian P-Spline Model to the data
We can then fit the Bayesian P-spline model to all of the REACT data for the first 7 rounds of the study (iterations and warmup will need to be bigger for accurate estimates)

```{r Fitting the p-spline model, results='hide'}
p_spline_mod_react <- stan_p_spline(pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r7,]$England,
                                   target_dist_between_knots = 5,
                                   spline_degree = 3,
                                   iter = 5000,
                                   warmup = 500,
                                   cores = 1)


```

The p-spline model can then be plotted with estimated 95%CI and 50%CI (the raw data used to produced the plots are avialble by using [[2]] and [[3]] instead of [[1]])
```{r Plotting the p-spline model}
p_spline_plot <- plot_p_spline_prev(pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$England,
                                   tot[tot$X>=min_date_r1 & tot$X<= max_date_r7,]$England,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   ylim = 2.0)

print(p_spline_plot[[1]])


```

From the p-spline model we can estimate the date of minimum prevalence and plot the posterior probability density

```{r Date of minimum from P-Spline model}
p_spline_min_date <- plot_p_spline_minimum_density(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3)

print(p_spline_min_date[[1]])


```
We can then calculate and plot the instanteous growth rate over the study period

```{r Instantaneous growth rate calculation}
p_spline_igr <- plot_p_spline_igr(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                  ylim = 0.2)

print(p_spline_igr[[1]])


```
and also the rolling two week average Reproduction number
```{r Reproduction number calculation}
p_spline_Rt <- plot_p_spline_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3)

print(p_spline_Rt[[1]])


```
From the Bayeisan P-spline posterior we can also calculate the average growth rate over periods of time covering individual and subsequent rounds.

```{r average Reproduction number calculation}
R_estimate_pspline_round1 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r1,
                                   max_date_num = max_date_r1,
                                   label = "Round1")
R_estimate_pspline_round2 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r2,
                                   max_date_num = max_date_r2,
                                   label = "Round2")
R_estimate_pspline_round3 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r3,
                                   max_date_num = max_date_r3,
                                   label = "Round3")
R_estimate_pspline_round4 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r4,
                                   max_date_num = max_date_r4,
                                   label = "Round4")
R_estimate_pspline_round5 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r5,
                                   max_date_num = max_date_r5,
                                   label = "Round5")
R_estimate_pspline_round6 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r6,
                                   max_date_num = max_date_r6,
                                   label = "Round6")
R_estimate_pspline_round7 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r7,
                                   max_date_num = max_date_r7,
                                   label = "Round7")

R_estimate_pspline_round12 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r1,
                                   max_date_num = max_date_r2,
                                   label = "Round12")
R_estimate_pspline_round23 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r2,
                                   max_date_num = max_date_r3,
                                   label = "Round23")
R_estimate_pspline_round34 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r3,
                                   max_date_num = max_date_r4,
                                   label = "Round34")
R_estimate_pspline_round45 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r4,
                                   max_date_num = max_date_r5,
                                   label = "Round45")
R_estimate_pspline_round56 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r5,
                                   max_date_num = max_date_r6,
                                   label = "Round56")
R_estimate_pspline_round67 <- estimate_p_spline_average_R(X = pos[pos$X>=min_date_r1 & pos$X<= max_date_r7,]$X,
                                   p_spline_fit = p_spline_mod_react, 
                                    target_dist_between_knots = 5,
                                    spline_degree = 3,
                                   min_date_num = min_date_r6,
                                   max_date_num = max_date_r7,
                                   label = "Round67")


R_table <- rbind(R_estimate_pspline_round1,R_estimate_pspline_round2,R_estimate_pspline_round3,
                 R_estimate_pspline_round4,R_estimate_pspline_round5,R_estimate_pspline_round6,
                 R_estimate_pspline_round7,
                 R_estimate_pspline_round12,R_estimate_pspline_round23,R_estimate_pspline_round34,
                 R_estimate_pspline_round45,R_estimate_pspline_round56,R_estimate_pspline_round67)
print(R_table)


```
